Repeat Yourself? When Linux kernel maintainers balance Code Duplication and Readability

### 2025-10-22 paper/techdebt2026/techdebt.tex
techdebt2025: adjustments (from the authors' meeting)
Signed-off-by: Paulo Meirelles <paulormm@ime.usp.br>

The detection of code duplication, or \textit{code clones}, has been a research topic for decades~\cite{firstman}. The literature offers a widely accepted taxonomy that classifies clones into four types based on their similarity, from exact copies (Type-1) to semantically equivalent but syntactically different fragments (Type-4)~\cite{litreview}. Several detection techniques have emerged, including textual, token-based, tree-based, and graph-based approaches~\cite{litreview}. These efforts have culminated in state-of-the-art techniques, such as the graph-based approach proposed by Liu et al.~\cite{tailor}.

@@

In the \textbf{complete participant observation} study, we chose two function pair found on AMD Display Driver to approach and apply the systematic approach, 
which we will denote as mitigation 1 and mitigation 2.

%TODO: ver se realmente queremos usar mitigation (ou contribution é melhor)

Table \ref{tab:patch} summarizes the impact of the mitigations in terms of 
files and lines changed and the refactoring method used. The following sections present our experience and learning for each mitigation.

\begin{table}[ht]
\centering
\caption{Refactor methods used and mitigation metrics in the participant-observation experiment.}
\begin{tabular}{ | m{12mm} | m{10mm} | m{10mm} | m{14mm} | m{20mm} | }

\hline

\textbf{Mitigation} & \textbf{Files changed} & \textbf{Lines added} & \textbf{Lines removed} & \textbf{Refactoring methods used}
\\ \hline

1 & 13 & +224 & -753 & Parameterize Method, Extract Method  \\ \hline
2 & 8 & +114 & -520 & None \\ \hline

\hline
\end{tabular}

\label{tab:patch}
\end{table}

@@

%TODO: deve ficar claro que isto aqui, mesmo que não dito explicitamente pelo mantenedores, é o resultado da Complete Participant Observation
\noindent\textbf{Takeaway 1:} Some of the code in the driver is shared across all operating systems that support the implemented GPUs. This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors. Additionally, \textbf{we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering}. One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others. This approach helps save a significant amount of time and effort.
\todo[inline]{...Foi tirado apenas das conversas com o Siqueira; tem o paper que o Gregório citou que menciona que duplicações podem ser bem vistas em drivers; isto um resultado/conclusão da Complete Participant Observation; e vamos citar que confirmamos o que está no paper Cloning Considered Harmful (Luan, acha que consegue fechar essa parte assim?)}

%TODO: ver se transicao estah boa; ou se eh melhor ter subsecoes
Regarding our analysis from the \textbf{participant-as-Observer} phase, Table~\ref{tab:stu} summarizes the student groups' mitigation approaches. It contains the information about the chosen subsystem and the similarity threshold for groups that opted for the first approach. Each entry also has the patch status, refactoring method, and code diff corresponding to the final patch versions before it was dropped or accepted into the code base. The code diff for a patchset corresponds to the sum of added and removed lines of each of its patches.

\begin{table}[ht]
\centering
\caption{Refactor methods used and metrics of the students' mitigations.}
\begin{tabular}{ |c |c |c |c |c | c| }
\hline
\textbf{G} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Method} & \textbf{Diff} \\
\hline

1 & IIO & 100\% & Dropped (v2) & Undefined & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & Parameterize & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & Parameterize & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & Undefined & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & Parameterize & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & Parameterize & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & Extract & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & Extract & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted* (v6) & Parameterize & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & Undefined & +32/-58 \\ \hline
11 & IIO & 100\% & Dropped (v1) & Inline & +2/-12 \\ 
   & AMD & 90\% & Accepted (v2) & Undefined & +90/-489 \\ \hline

@@

Seven groups applied the parameterized refactor method, two groups used the extractor method, and one group applied the inline method to their approaches to mitigate their duplications. Four patches do not contain a refactoring technique defined in the literature (undefined). Six groups had their patches accepted and fully merged into Linux; four groups had their patches dropped; one group has its patchset partially applied and is still working on patch submission, currently in the sixth version of the patch when this document was written, but we can also consider this case as accepted.

>>>>

%TODO: temos que citar o conceito de DRY aqui ... (Book: Pragmatic Programmer, The: From Journeyman to Master)
% We feel that the only way to develop software reliably, and to make our developments easier to understand and maintain, is to follow what we call the DRY principle: Every piece of knowledge must have a single, unambiguous, authoritative representation within a system.
The detection of code duplication, or \textit{code clones}, has been a research topic for decades~\cite{firstman}. The literature offers a widely accepted taxonomy that classifies clones into four types based on their similarity, from exact copies (Type-1) to semantically equivalent but syntactically different fragments (Type-4)~\cite{litreview}. Several detection techniques have emerged, including textual, token-based, tree-based, and graph-based approaches~\cite{litreview}. These efforts have culminated in state-of-the-art techniques, such as the graph-based approach proposed by Liu et al.~\cite{tailor}.

@@

In the \textbf{complete participant observation} study, we chose two function pair found on AMD Display Driver to approach and apply the systematic approach ...

%TODO (Luan): nao terah mitigation 1 e 2, apenas a contribuicao enviada mesmo.

@@

%TODO: deve ficar claro que isto aqui, mesmo que não dito explicitamente pelo mantenedores, é o resultado da Complete Participant Observation
\noindent\textbf{Takeaway T1:} Some of the code in the driver is shared across all operating systems that support the implemented GPUs. This fact creates a complex process within AMD to format and submit changes across the supported systems while simultaneously implementing measures to mitigate errors. Additionally, \textbf{we comprehended that AMD Display driver developers can not view duplicated code negatively, as we understand it is a good practice in software engineering}. One example shared with us is that duplicated code enhances the independence of GPU driver code, allowing developers to make changes to a specific GPU without needing to test compatibility with others. This approach helps save a significant amount of time and effort.
\todo[inline]{...Foi tirado apenas das conversas com o Siqueira; tem o paper que o Gregório citou que menciona que duplicações podem ser bem vistas em drivers; isto um resultado/conclusão da Complete Participant Observation; e vamos citar que confirmamos o que está no paper Cloning Considered Harmful (Luan, acha que consegue fechar essa parte assim?)}

%TODO: ver se transicao estah boa; ou se eh melhor ter subsecoes
Regarding our analysis from the \textbf{participant-as-Observer} phase, Table~\ref{tab:stu} summarizes the student groups' mitigation approaches. It contains the information about the chosen subsystem (SS) and the similarity threshold for groups (SML) that opted for the first approach. Each entry also has the patch status, and code diff corresponding to the final patch versions before it was dropped or accepted into the code base. The code diff for a patchset corresponds to the sum of added and removed lines of each of its patches.

\begin{table}[ht]
\centering
\caption{Refactor methods used and metrics of the students' mitigations.}
\begin{tabular}{ |c |c |c |c |c | c| }
\hline
\textbf{GID} & \textbf{SS} & \textbf{SML} & \textbf{Patch Status} & \textbf{Takeaway} & \textbf{Diff} \\
\hline
0 & AMD & ???\% & Accepted (v3) & T1 & +114/-520 \\ \hline
1 & IIO & 100\% & Dropped (v2) & ? & +7/-14 \\ \hline
2 & IIO & 100\% & Dropped (v1) & ? & +23/-27 \\ \hline
3 & IIO & 100\% & Accepted (v1) & ? & +12/-30 \\ \hline
4 & IIO & 100\% & Accepted (v3) & ? & +2/-7 \\ \hline
5 & IIO & 90\% & Accepted (v1) & ? & +17/-70 \\ \hline
6 & IIO & 90\% & Dropped (v1) & ? & +95/-92 \\ \hline
7 & IIO & 90\% & Accepted (v3) & ? & +20/-36 \\ \hline
8 & IIO & 90\% & Accepted (v4) & ? & +20/-16 \\ \hline
9 & IIO & 90\% & Accepted (v6) & ? & +149/-243 \\ \hline
10 & IIO & 90\% & Dropped (v1) & ? & +32/-58 \\ \hline
11 & IIO & 100\% & Dropped (v1) & ? & +2/-12 \\ 
   & AMD & 90\% & Accepted (v2) & ? & +90/-489 \\ \hline

@@

Regarding to refactor methods ....

%TODO: nao descrever e discutir mais as estrategias de refatoracao por patch ...
Seven groups applied the parameterized refactor method, two groups used the extractor method, and one group applied the inline method to their approaches to mitigate their duplications. Four patches do not contain a refactoring technique defined in the literature (undefined). Six groups had their patches accepted and fully merged into Linux; four groups had their patches dropped; one group has its patchset partially applied and is still working on patch submission, currently in the sixth version of the patch when this document was written, but we can also consider this case as accepted.
